Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-12 paper/paper.tex
more and updated results

\begin{table}[ht]
\centering
\begin{tabular}{lr}
  \toprule
& mean runtime\\
  \midrule
virtual best & 5822.809\\
  single best & 7788.719\\
  LLAMA & 6509.947\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance on reduced set of 2336
instances.}\label{tab:res}
\end{table}

We are able to close more than 60\% of the performance gap between the single
best and the virtual best solver, our lower and upper bounds. For comparison, we
show the results achieved on the entire set of instances without
unsatisfiability detection and presolving in Table~\ref{tab:resfull}. We achieve
similar results with more than 58\% of the gap between the single best and the
virtual best closed.

\begin{table}[ht]
\centering
\begin{tabular}{lr}
  \toprule
& mean runtime\\
  \midrule
virtual best & 2375.913\\
  single best & 3177.569\\
  LLAMA & 2706.749\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance on the full set of 5725
instances.}\label{tab:resfull}
\end{table}

Figure~\ref{fig:scatter} shows detailed results for each of the 2336 instances
in the reduced set. Our algorithm selection approach is by no means able to
identify the best solver on a majority of instances, although we are able to
achieve performance improvements overall. There seems to be no particular
pattern as to where algorithm selection performs well and where it fails.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/perfScatter}
\caption{Single best algorithm vs.\ portfolio performance on all instances by
family. Points above the diagonal indicate that the portfolio achieved better
performance, while the single best solver was better on instances below the
diagonal.}
\label{fig:scatter}
\end{figure}

Figure~\ref{fig:portfolio-ecdf} shows the cumulative distribution function for
the individual solvers, the virtual best solver, and the LLAMA portfolio. The
virtual best solver clearly dominates with a significant margin to the next
best. VF2 is far below all other solvers. The portfolio does not perform well
for instances that can be solved quickly because of the overhead incurred
through feature computation. As the instances become more difficult to solve,
its performance improves. For instances that take more than 1,000 seconds to
solve, the portfolio eclipses any individual solver and closes most of the gap
to the virtual best solver.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
\caption{Empirical cumulative distribution function of solvers, virtual best
solvers, and the LLAMA portfolio.}
\label{fig:portfolio-ecdf}
\end{figure}

>>>>

\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \toprule
model & mean MCP & solved instances & mean performance\\
  \midrule
virtual best & 0 & 2219 & 5822809\\
single best & 1965158 & 2172 & 7787967\\
LLAMA & 610835 & 2206 & 6435301\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance on the reduced set of 2336
instances. MCP is the misclassification penalty; that is, the additional time
required to solve a set of instances because of choosing solvers that perform
worse than the best.}\label{tab:res} \end{table}

We are able to solve 34 more instances than the single best solver within the
timeout, with only an additional 13 to the virtual best. In terms of average
performance, we are able to close almost 70\% of the gap between the single best
and the virtual best solver.

Table~\ref{tab:res} shows the performance of the selection model. The
performance of the entire system we propose here is shown in
Table~\ref{tab:resfull}. The additional instances are solved either by
incompleteLAD or VF2 presolving, as explained above. Our system is able to close
more than 50\% of the gap between single and virtual best.

\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \toprule
model & mean MCP & solved instances & mean performance\\
  \midrule
virtual best & 0 & 5608 & 2375913\\
single best & 798660 & 5562 & 3174573\\
LLAMA & 249242 & 5595 & 2625831\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection system performance on the full set of 5725
instances.}\label{tab:resfull}
\end{table}

Figure~\ref{fig:portfolio-ecdf} shows the cumulative distribution function for
the individual solvers, the virtual best solver, and the LLAMA portfolio. The
virtual best solver clearly dominates with a significant margin to the next
best. VF2 is far below all other solvers. The portfolio does not perform well
for instances that can be solved quickly because of the overhead incurred
through feature computation. As the instances become more difficult to solve,
its performance improves.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
\caption{Empirical cumulative distribution function of solvers, virtual best
solvers, and the LLAMA portfolio on the reduced set of 2336 instances. The
actual solvers are shown as dashed lines.}
\label{fig:portfolio-ecdf}
\end{figure}

Figure~\ref{fig:portfolio-ecdf-full} shows the ECDF of the system including
incompleteLAD and VF2 presolving on the full set of instances. Note that for
very easy instances, we are better than the virtual best solver because
incompleteLAD is not included in our portfolio. The performance on small
instances is much better than the LLAMA selector alone (cf.\
Figure~\ref{fig:portfolio-ecdf}) and the region where the portfolio performs
worse than the individual solvers is now limited to approximately $10^2$ to
$10^4$ milliseconds.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/portfolio-ecdf-full}
\caption{Empirical cumulative distribution function of solvers, virtual best
solvers, and the portfolio system on the full set of 5725 instances. The
actual solvers are shown as dashed lines.}
\label{fig:portfolio-ecdf-full}
\end{figure}

We train the algorithm selection model specifically for the timeout of $10^8$
milliseconds. In particular, we are interested in minimising the performance
difference to the virtual best. Problem instances that take longer to solver
contribute more to this difference than easy instances and therefore carry more
weight for the algorithm selection model. That is, choosing the wrong solver for
a hard instances is much worse than choosing the wrong solver for an easy
instance.

Even though Figures~\ref{fig:portfolio-ecdf} and~\ref{fig:portfolio-ecdf-full}
show that on easy instances, the portfolio does not perform well, it does over
the entire set of instances (cf.\ Tables~\ref{tab:res} and~\ref{tab:resfull}).

\subsection{Portfolios for different timeouts}

The question arises of whether the success of our approach is limited to the
particular timeout we have chosen. Table~\ref{tab:resTimeouts} shows the
performance of our portfolio for timeouts ranging from $10^2$ to $10^8$. We
do not include incompleteLAD and VF2 presolving here, as these components do not
change for different timeouts.

\begin{table}[ht]
\centering
\begin{tabular}{llrrr}
  \toprule
timeout & solver & mean MCP & solved instances & mean performance\\
  \midrule
$10^2$ & virtual best & 0 & 1253 & 59\\
 & LLAMA & 3 & 685 & 78\\
 & glasgow1 & 11 & 903 & 71\\
  \midrule
$10^3$ & virtual best & 0 & 1781 & 322\\
 & LLAMA & 22 & 1284 & 569\\
 & glasgow1 & 126 & 1539 & 449\\
  \midrule
$10^4$ & virtual best & 0 & 1943 & 2018\\
 & LLAMA & 186 & 1771 & 3342\\
 & glasgow2 & 683 & 1855 & 2701\\
  \midrule
$10^5$ & virtual best & 0 & 2045 & 14348\\
 & LLAMA & 1346 & 2022 & 17231\\
 & glasgow2 & 3019 & 1988 & 17368\\
  \midrule
$10^6$ & virtual best & 0 & 2111 & 108060\\
 & LLAMA & 10826 & 2087 & 120459\\
 & glasgow2 & 22729 & 2063 & 130789\\
  \midrule
$10^7$ & virtual best & 0 & 2178 & 815577\\
 & LLAMA & 91012 & 2156 & 908200\\
 & glasgow2 & 198944 & 2129 & 1014522\\
  \midrule
$10^8$ & virtual best & 0 & 2219 & 5822809\\
 & LLAMA & 740765 & 2204 & 6565232\\
 & glasgow2 & 1969418 & 2172 & 7792227\\
\bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance for different timeouts.}\label{tab:resTimeouts}
\end{table}

The results show that while for the misclassification penalty, LLAMA is always
better than the single best algorithm, for timeouts $10^2$ to $10^4$ it is worse
in terms of the number of solved instances and mean performance. An explanation
for this is that empirical runtime measurements are inherently noisy, which
affects the accuracy of the measurements particularly in this range.
